---
title: "Extracting purse-seine catch and length compostion data for bigeye  in 1975-1999"
author: "Haikun Xu"
date: "`r Sys.Date()`"
output:
  word_document:
    fig_width: 8
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = FALSE)
```

This example code demonstrates how to extract the purse-seine catch and length composition data for the stock assessment of bigeye tuna in the eastern Pacific Ocean. Data are extracted for bigeye between 1975 and 1999 based on the R package *BSE* (version `r packageVersion("BSE")`). The package can be installed using `devtools::install_github('HaikunXu/BSE',ref='main')`. Fishery definition for this data extraction is based on the benchmark assessment conducted in 2020.

-   Step 1: set up some directories and parameters for the extraction

```{r global setup}
# devtools::install_github('HaikunXu/BSE',ref='main') 
library(BSE)

# Load the base files (please ask Haikun to get those data)
load("D:/OneDrive - IATTC/IATTC/2022/BSE stuff from Cleridy/single spp programs_PS_1975-1999/Raw data extractions/base files_1975-1999_for SAC 2020.RData")
# the directory where output will be saved
save_dir <- "D:/OneDrive - IATTC/IATTC/2022/BSE stuff from Cleridy/BET/"
yr.start <- 1975
yr.end <- 1999
Species <- "BET"
grow.increments <- grow.increments.betyftskj # the growth increment matrix
```

-   Step 2: compile the OBJ catch and composition data for BET

```{r OBJ}
PS <- "OBJ"
cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
```

Check the strata definition for OBJ to make sure that it is correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for BET in the OBJ fishery

```{r}
for(year in yr.start:yr.end) {
  #print(paste0("Year: ",year))
  
  #print("Step 1: get well estimates")
  well.estimates <- well.estimates.7599.f(year)
  
  #print("Step 2: get catch estimates")
  if(year!=1988) catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,2,well.estimates,PS,Species)
  else catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,1,well.estimates,PS,Species)
  
  #print("Step 3: get fishery estimates")
  fishery.estimates <- call.fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$totunlds.bystrat,year,PS,Species)
  
  fishery.estimates.bet <- fishery.estimates$bet
  
  assign(paste0("fishery.estimates.bet.", year), fishery.estimates.bet, pos=1)
}
save(list=objects(pat="fishery.estimates.bet."),file=paste0(save_dir,"BET_",PS,"_1975-1999.RData"))
```

Get final OBJ catch and comp output for the stock assessment

```{r}
BET.OBJ.Catch.19751999<-compile.catch.output.7599.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2","A3","A4","A5"))
BET.OBJ.Comp.19751999<-compile.sizecomps.output.7599.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 3: compile the NOA catch and composition data for BET

```{r NOA}
PS <- "NOA"
cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
```

Check the strata definition for NOA to make sure that it is correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for BET in the NOA fishery

```{r}
for(year in yr.start:yr.end) {
  #print(paste0("Year: ",year))
  
  print("Step 1: get well estimates")
  well.estimates <- well.estimates.7599.f(year)
  
  #print("Step 2: get catch estimates")
  if(year!=1988) catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,2,well.estimates,PS,Species)
  else catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,1,well.estimates,PS,Species)
  
  #print("Step 3: get fishery estimates")
  fishery.estimates <- call.fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$totunlds.bystrat,year,PS,Species)
  
  fishery.estimates.bet <- fishery.estimates$bet
  
  assign(paste0("fishery.estimates.bet.", year), fishery.estimates.bet, pos=1)
}
save(list=objects(pat="fishery.estimates.bet."),file=paste0(save_dir,"BET_",PS,"_1975-1999.RData"))
```

Get final NOA catch and comp output for the stock assessment

```{r}
BET.NOA.Catch.19751999<-compile.catch.output.7599.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2"))
BET.NOA.Comp.19751999<-compile.sizecomps.output.7599.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 4: compile the DEL catch and composition data for BET

```{r DEL}
PS <- "DEL"
cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
```

Check the strata definition for DEL to make sure that it is correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for BET in the DEL fishery

```{r}
for(year in yr.start:yr.end) {
  #print(paste0("Year: ",year))
  
  
  #print("Step 1: get well estimates")
  well.estimates <- well.estimates.7599.f(year)
  
  #print("Step 2: get catch estimates")
  if(year!=1988) catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,2,well.estimates,PS,Species)
  else catch.estimates <- get.catch.estimates.7599.f(cae,cae.stratflg,corrected.unlds,lfgrpd,lfmm,year,1,well.estimates,PS,Species)
  
  #print("Step 3: get fishery estimates")
  fishery.estimates <- call.fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$totunlds.bystrat,year,PS,Species)
  
  fishery.estimates.bet <- fishery.estimates$bet
  
  assign(paste0("fishery.estimates.bet.", year), fishery.estimates.bet, pos=1)
}
save(list=objects(pat="fishery.estimates.bet."),file=paste0(save_dir,"BET_",PS,"_1975-1999.RData"))
```

Get final DEL catch and comp output for the stock assessment

```{r}
BET.DEL.Catch.19751999<-compile.catch.output.7599.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2"))
BET.DEL.Comp.19751999<-compile.sizecomps.output.7599.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 6: save all results for BET as csv files

```{r save results}
write.csv(BET.OBJ.Catch.19751999,file=paste0(save_dir,"BET.OBJ.Catch.19751999.csv"),row.names = FALSE)
write.csv(BET.OBJ.Comp.19751999,file=paste0(save_dir,"BET.OBJ.Comp.19751999.csv"),row.names = FALSE)
write.csv(BET.NOA.Catch.19751999,file=paste0(save_dir,"BET.NOA.Catch.19751999.csv"),row.names = FALSE)
write.csv(BET.NOA.Comp.19751999,file=paste0(save_dir,"BET.NOA.Comp.19751999.csv"),row.names = FALSE)
write.csv(BET.DEL.Catch.19751999,file=paste0(save_dir,"BET.DEL.Catch.19751999.csv"),row.names = FALSE)
write.csv(BET.DEL.Comp.19751999,file=paste0(save_dir,"BET.DEL.Comp.19751999.csv"),row.names = FALSE)
```
