---
title: "Extracting purse-seine catch and length compostion data for yellowfin in 2000-2022"
author: "Haikun Xu"
date: "`r Sys.Date()`"
output:
  word_document:
    fig_width: 8
    fig_height: 6
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = FALSE)
```

This example code demonstrates how to extract the purse-seine catch and length composition data for the stock assessment of yellowfin tuna in the eastern Pacific Ocean. Data are extracted for yellowfin between 2000 and 2022 based on the R package *BSE* (version `r packageVersion("BSE")`). The package can be installed using `devtools::install_github('HaikunXu/BSE',ref='main')`. Fishery definition for this data extraction is based on the benchmark assessment conducted in 2020.

-   Step 1: set up some directories and parameters for the extraction

```{r global setup}
# devtools::install_github('HaikunXu/BSE',ref='main') 
library(BSE)
library(tidyverse)

# Load the base files (please ask Haikun to get those data)
raw_data_dir <- "D:/OneDrive - IATTC/IATTC/2024/SAC15/PS Database/"
# the directory where output will be saved
save_dir <- "D:/OneDrive - IATTC/IATTC/2022/BSE stuff from Cleridy/YFT/"
yr.start <- 2000
yr.end <- 2023
Species <- "YFT"
grow.increments <- grow.increments.betyftskj # the growth increment matrix

area.substitution.mat.OBJ <- matrix(
  c(1, 2, 4, 3, 5,
    2, 1, 4, 3, 5,
    3, 4, 2, 1, 5,
    4, 3, 2, 5, 1,
    5, 4, 3, 2, 1),
  ncol = 5,
  byrow = TRUE
)

area.substitution.mat.NOA <- matrix(
  c(1, 2, 4, 3, 5,
    2, 1, 4, 3, 5,
    3, 4, 2, 1, 5,
    4, 3, 2, 5, 1,
    5, 4, 3, 2, 1),
  ncol = 5,
  byrow = TRUE
)

area.substitution.mat.DEL <- matrix(
  c(1, 2, 4, 3,
    2, 1, 4, 3,
    3, 4, 2, 1,
    4, 3, 2, 1),
  ncol = 4,
  byrow = TRUE
)

#fishery substitute matrix
my.FOmatrix <- matrix(paste0("FO.A", area.substitution.mat.OBJ),ncol=ncol(area.substitution.mat.OBJ))

my.UNmatrix <- matrix(paste0("UN.A", area.substitution.mat.NOA),ncol=ncol(area.substitution.mat.NOA))

my.DPmatrix <- matrix(paste0("DP.A", area.substitution.mat.DEL),ncol=ncol(area.substitution.mat.DEL))
```

-   Step 2: lightly process the raw data so that they can be easily used in rest of steps

```{r process raw data}
# Get the total unloads for the PS fleet
total.unlds <- read.unloads.f(raw_data_dir,"Unloading2000-2023.txt",yr.start,yr.end)
# Get the CAE+IDM data
cae <- read.cae.f(raw_data_dir,"CAE-LatLon2000-2023.txt",yr.start,yr.end)
# Get the length-frequency data (length in millimeters)
lfmm <- read.lfmmdata.f(raw_data_dir,"LengthMM2000-2023.txt")
# Get the grouped length-frequency output
lfgrpd <- read.lengthfreq.f(raw_data_dir,"LengthFreq2000-2023.txt")
```

-   Step 3: compile the OBJ catch and composition data for YFT

```{r OBJ}
PS <- "OBJ"
area.substitution.mat <- area.substitution.mat.OBJ

cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
lfgrpd.stratflg <- create.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,is.lwrght=T,floor(lfgrpd$moda/100),lfgrpd$setype,lfgrpd$class,PS=PS,Species=Species)
```

Check the strata definition for OBJ in both cae and lf data sets to make sure that they are correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
# check.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,lfgrpd.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for YFT in the OBJ fishery

```{r}
for(year in yr.start:yr.end) {
  print(paste0("Year: ",year))
  
  # print("Step 1: get well estimates")
  well.estimates <- well.estimates.f(lfgrpd[lfgrpd$year.firstset==year,],lfmm)
  
  # print("Step 2: get catch estimates")
  catch.estimates <- get.catch.estimates.f(cae,cae.stratflg,total.unlds,lfgrpd,lfgrpd.stratflg,lfmm,year,2,well.estimates,area.substitution.mat,grow.increments,PS=PS,Species=Species,my.FOmatrix,my.UNmatrix,my.DPmatrix)
  
  # print("Step 3: get fishery estimates")
  fishery.estimates <- fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$stratum.estimates.NOsamps,year,PS=PS,Species=Species)
  
  assign(paste0("fishery.estimates.", year), fishery.estimates, pos=1)
}
# save middle-step data as a record
save(list=objects(pat="fishery.estimates"),file=paste0(save_dir,"YFT_",PS,"_2000-2023.RData"))
```

Get final OBJ catch and comp output for the stock assessment

```{r}
YFT.OBJ.Catch.20002023<-compile.catch.output.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2","A3","A4","A5")) # five OBJ fisheries
YFT.OBJ.Comp.20002023<-compile.sizecomps.output.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 4: compile the NOA catch and composition data for YFT

```{r NOA}
PS <- "NOA"
area.substitution.mat <- area.substitution.mat.NOA

cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
lfgrpd.stratflg <- create.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,is.lwrght=T,floor(lfgrpd$moda/100),lfgrpd$setype,lfgrpd$class,PS=PS,Species=Species)
```

Check the strata definition for NOA in both cae and lf data sets to make sure that they are correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
check.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,lfgrpd.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for YFT in the NOA fishery

```{r}
for(year in yr.start:yr.end) {
  print(paste0("Year: ",year))
  
  # print("Step 1: get well estimates")
  well.estimates <- well.estimates.f(lfgrpd[lfgrpd$year.firstset==year,],lfmm)
  
  # print("Step 2: get catch estimates")
  catch.estimates <- get.catch.estimates.f(cae,cae.stratflg,total.unlds,lfgrpd,lfgrpd.stratflg,lfmm,year,2,well.estimates,area.substitution.mat,grow.increments,PS=PS,Species=Species,my.FOmatrix,my.UNmatrix,my.DPmatrix)
  
  # print("Step 3: get fishery estimates")
  fishery.estimates <- fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$stratum.estimates.NOsamps,year,PS=PS,Species=Species)
  
  assign(paste0("fishery.estimates.", year), fishery.estimates, pos=1)
}
# save middle-step data as a record
save(list=objects(pat="fishery.estimates"),file=paste0(save_dir,"YFT_",PS,"_2000-2023.RData"))
```

Get final NOA catch and comp output for the stock assessment

```{r}
YFT.NOA.Catch.20002023<-compile.catch.output.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2","A3","A4","A5"))  # four NOA fisheries
YFT.NOA.Comp.20002023<-compile.sizecomps.output.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 5: compile the DEL catch and composition data for YFT

```{r DEL}
PS <- "DEL"
area.substitution.mat <- area.substitution.mat.DEL

cae.stratflg <- create.strat.flg.f(cae$latc5,cae$lonc5,is.lwrght=F,cae$month,cae$setype,cae$class,PS=PS,Species=Species)
lfgrpd.stratflg <- create.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,is.lwrght=T,floor(lfgrpd$moda/100),lfgrpd$setype,lfgrpd$class,PS=PS,Species=Species)
```

Check the strata definition for DEL in both cae and lf data sets to make sure that they are correct

```{r}
check.strat.flg.f(cae$latc5,cae$lonc5,cae.stratflg)
# check.strat.flg.f(lfgrpd$lat.5deg,lfgrpd$lon.5deg,lfgrpd.stratflg)
```

Loop through every year between `yr.start` and `yr.end` to get catch and composition data for YFT in the DEL fishery

```{r}
for(year in yr.start:yr.end) {
  print(paste0("Year: ",year))
  
  # print("Step 1: get well estimates")
  well.estimates <- well.estimates.f(lfgrpd[lfgrpd$year.firstset==year,],lfmm)
  
  # print("Step 2: get catch estimates")
  catch.estimates <- get.catch.estimates.f(cae,cae.stratflg,total.unlds,lfgrpd,lfgrpd.stratflg,lfmm,year,2,well.estimates,area.substitution.mat,grow.increments,PS=PS,Species=Species,my.FOmatrix,my.UNmatrix,my.DPmatrix)
  
  # print("Step 3: get fishery estimates")
  fishery.estimates <- fishery.estimates.f(catch.estimates$stratum.estimates.withsamps,catch.estimates$stratum.estimates.NOsamps,year,PS=PS,Species=Species)
  
  assign(paste0("fishery.estimates.", year), fishery.estimates, pos=1)
}
# save middle-step data as a record
save(list=objects(pat="fishery.estimates"),file=paste0(save_dir,"YFT_",PS,"_2000-2023.RData"))
```

Get final DEL catch and comp output for the stock assessment

```{r}
YFT.DEL.Catch.20002023<-compile.catch.output.f(yr.start,yr.end,PS=PS,Species=Species,c("A1","A2","A3","A4")) # two DEL fisheries
YFT.DEL.Comp.20002023<-compile.sizecomps.output.f(yr.start,yr.end,PS=PS,Species=Species)
```

-   Step 6: save all results for YFT as csv files

```{r save results}
write.csv(YFT.OBJ.Catch.20002023,file=paste0(save_dir,"YFT.OBJ.Catch.20002023.csv"),row.names = FALSE)
write.csv(YFT.OBJ.Comp.20002023,file=paste0(save_dir,"YFT.OBJ.Comp.20002023.csv"),row.names = FALSE)
write.csv(YFT.NOA.Catch.20002023,file=paste0(save_dir,"YFT.NOA.Catch.20002023.csv"),row.names = FALSE)
write.csv(YFT.NOA.Comp.20002023,file=paste0(save_dir,"YFT.NOA.Comp.20002023.csv"),row.names = FALSE)
write.csv(YFT.DEL.Catch.20002023,file=paste0(save_dir,"YFT.DEL.Catch.20002023.csv"),row.names = FALSE)
write.csv(YFT.DEL.Comp.20002023,file=paste0(save_dir,"YFT.DEL.Comp.20002023.csv"),row.names = FALSE)
```